---
title: "Lab-11 Replication File"
output: html_notebook
---

## Introduction
This notebook replicates the lab for week 11.

## Dependencies
This notebook requires the following packages:

```{r}
library(effsize)
library(dplyr)
library(ggplot2)
library(psych)
library(pwr)
library(stargazer)
library(testDriveR)
```

We also use the following custom function for creating correlation tables:

```{r}
corrTable <-function(.data, coef = c("pearson", "spearman"), listwise = TRUE, round = 3, pStar = TRUE, ...){

  ## process dots
  if (rlang::dots_n(...) > 1) {
    .data <- dplyr::select(.data, ...)
  }
  
  ## listwise deletion
  if (listwise == TRUE) {
    .data <- na.omit(.data)
  }
  
  ## compute correlation matrix
  inputMatrix <- as.matrix(.data)
  corrMatrix <- Hmisc::rcorr(inputMatrix, type = coef)
  
  ## matrix of correlation coeficients
  rCoef <- corrMatrix$r
  
  # matrix of p-values
  pValues <- corrMatrix$P  
  
  ## round the correlation matrix values
  rCoef <- format(round(cbind(rep(-1.11, ncol(.data)), rCoef), round))[,-1]
  
  ## statistical significance stars
  if (pStar == TRUE) {
    
    ## Define notions for significance levels
    stars <- ifelse(pValues < .001, "***", 
                    ifelse(pValues < .01, "**", 
                           ifelse(pValues < .05, "*", "")))
    
    ## add apropriate stars
    rCoef <- matrix(paste(rCoef, stars, sep = ""), ncol = ncol(.data))    
  }
  
  ## remove upper triangle
  rCoef[upper.tri(rCoef, diag = FALSE)] <- ""
  diag(rCoef) <- formatC(1, digits = round, format = "f")
  
  ## final formatting
  rCoef <- as.data.frame(rCoef)
  rCoef <- dplyr::mutate_if(rCoef, is.factor, as.character)
  rownames(rCoef) <- colnames(.data)
  colnames(rCoef) <- paste(colnames(.data), "", sep = "")
  
  ## return data frame
  return(rCoef)
} 
```

## Data
We'll assign the data for the lab to a new object named `autoData`.

```{r}
autoData <- auto17
```

## Part 1: Correlation in `R`
### Question 1
The following syntax creates a correlation table for the variables `combFE`, `fuelCost`, and `cyl`.

```{r}
table <- corrTable(autoData, coef = "pearson", listwise = TRUE, round = 3, pStar = TRUE, combFE, fuelCost, cyl)
table
```

### Question 2
There is a strong, negative relationship between combined fuel efficiency and fuel cost ($r = -.899, p < .001$). The higher the fuel cost, the lower the fuel efficiency of the vehicle.

### Question 3
There is a strong, positive relationship between fuel cost and number of cylinders ($r = .843, p < .001$). The higher the number of cylinders in the engine, the higher the fuel cost of the vehicle. 

### Question 4
There is a strong, negative relationship between combined fuel efficiency and number of cylinders ($r = -.723, p < .001$). The higher the number of cylinders in the engine, the lower the fuel efficiency of the vehicle.

### Question 5
The following code creates a LaTeX formatted table

```{r}
stargazer(table, title = "Correlation Table for Vehicle Efficiency", summary = FALSE)
```

## Part 2: Scatterplots
### Question 6
The following code produces a scatterplot between the variables `combFE` and `fuelCost`.

```{r}
ggplot(data = autoData) +
  geom_point(mapping = aes(x = combFE, y = fuelCost), position = "jitter")
```

We can see a clear, if non-linear, relationship between the two variables. As fuel efficiency increases, the cost of fuel declines.

### Question 7

The following code creates a binary variable representing gas guzzling vehicles (as defined by the federal Department of Transportation).

```{r}
autoData <- mutate(autoData, guzzlerLog = ifelse(guzzlerStr == "G", TRUE, FALSE))
```

### Question 8

We can use the new guzzler variable to color the points of our scatterplot.

```{r}
ggplot(data = autoData) +
  geom_point(mapping = aes(x = combFE, y = fuelCost, color = guzzlerLog), position = "jitter")
```

This plot shows that all of the guzzler vehicles have both high fuel cost and lost fuel efficiency.

### Question 9

We can also use the new guzzler variable as a faceting variable to more clearly highlight the pattern unique to guzzlers.

```{r}
ggplot(data = autoData) +
  geom_point(mapping = aes(x = combFE, y = fuelCost, color = guzzlerLog), position = "jitter") +
  facet_wrap( ~ guzzlerLog)
```

### Question 10
This produces an error - apologies!

## Part 3: Cronbach's Alpha
### Question 11
The following code creates a subset of the `autoData` data frame, and then calculates a Cronbach's alpha value from it. 

```{r}
autoData %>%
  select(combFE, fuelCost, cyl) -> autoSelection

psych::alpha(autoSelection, check.keys = TRUE)
```

The alpha value ($alpha = .93$) is particularly high, suggesting that there may be an underlying construct (engine efficiency) that all three of these variables measure.

## Part 4: Sample Size Estimates
## Question 12
The following code calculates an estimated sample size for the given parameters.

```{r}
pwr.r.test(r = .89, sig.level = .05, power = .8, alternative = "two.sided")
```

We would need a sample of at least 7 people (in reality, 30) to see this effect size ($r=.89$) at the given level of power ($power = .8$)

## Question 13
The following code calculates an estimated sample size for the given parameters.

```{r}
pwr.r.test(r = .29, sig.level = .05, power = .9, alternative = "two.sided")
```

We would need a sample of at least 120 people to see this effect size ($r=.29$) at the given level of power ($power = .9$)
